Deep mutational scanning of hemagglutinin helps predict the evolutionary fates of human H3N2 influenza virus variants

### 2018-03-29 paper/manuscript.tex
elaborated on comments to @jlhudd in manuscript text

\begin{figure}
\centering
\includegraphics[width=\linewidth]{figs/frequency_trajectory_ex.png}
\caption{\label{fig:frequency_trajectory}
{\bf Frequency trajectories of viral clades and their relation to the effects of individual mutations.}
The top panel shows the subset of the full H3N2 HA tree in Figure~\ref{suppfig:tree} windowed from 2004 to 2014. 
Circles indicate amino-acid mutations, and are colored according to the mutational effect measured in our deep mutational scanning (negative values indicate mutations measured to be deleterious to viral growth).
The bottom panel shows the frequency trajectory of the clade downstream of each mutation, with clades colored according to the mutational effects as in the top panel.
It is obvious that most clades that reach high frequency have mutations that are measured to be relatively favorable in our experiments.
\comment{John: Can you update figure? In particular, maybe better colors. Also, let's increase the size of the legend and make it $<-10$, and $> 0$ rather than having the \emph{inf} in the labels. Also, can you make y-axis at bottom ``clade frequency.''}
}
\end{figure}

@@

\begin{figure*}[ht]
\centering
\includegraphics[width=\textwidth]{figs/muteffect_by_maxfreq.pdf}
\caption{\label{fig:muteffect_maxfreq}
{\bf The experimental measurements are informative about the evolutionary fate of viral clades.}
(A) The correlation between the effects of mutations as measured in our experiments and the maximum frequency reached by viral clades.
The plots show the Spearman correlation $\rho$ and a $P$-value obtained by repeatedly randomizing the experimental measurements among sites \comment{John: is this right about how the $P$-value was calculated?}.
(B) The distribution of mutational effects among clades binned by the maximum frequency that they reach.
Note that the analysis is performed separately for the pre- and post-Perth/2009 partitions of the tree (see Figure~\ref{suppfig:tree}).
\comment{Split the top and bottom into A and B, adjust font sizes (particularly, tick labels too big), adjust capitalization to be consistent with other figures (which usually start lower-case unless proper noun), increasing spacing between A and B panels. Also, no need for a title at top and figure can probably get smaller.}
}
\end{figure*}

@@

\begin{figure*}[ht]
\centering
\includegraphics[width=\textwidth]{figs/muteffect_by_maxfreq_WSN_1.pdf}
\caption{\label{fig:muteffect_maxfreq_WSN}
{\bf Experimental measurements on an H1 HA are not informative about the evolutionary fate of H3N2 viral clades.}
This figure repeats the analysis of the H3N2 clade frequencies in Figure~\ref{fig:muteffect_maxfreq}A, but uses the deep mutational scanning data for an H1 HA as measured in \cite{doud2016accurate}.
Figure~\ref{suppfig:muteffect_maxfreq_WSN_supp} shows the histograms comparable to those in Figure~\ref{fig:muteffect_maxfreq}B.
}
\end{figure*}

@@

\begin{suppfigure}[H]
\centerline{\includegraphics[width=0.75\textwidth]{figs/muteffect_by_maxfreq_WSN_2.pdf}}
\caption{\label{suppfig:muteffect_maxfreq_WSN_supp}
{\bf The distribution of mutational effects measured in H1 HA among H3N2 clades binned by the maximum frequency that they reach.}
This figure repeats the analysis of the H3N2 clade frequencies in Figure~\ref{fig:muteffect_maxfreq}B, but uses the deep mutational scanning data for an H1 HA as measured in \cite{doud2016accurate}.
}
\end{suppfigure}

>>>>

\begin{figure}
\centering
\includegraphics[width=\linewidth]{figs/frequency_trajectory_ex.png}
\caption{\label{fig:frequency_trajectory}
{\bf Frequency trajectories of viral clades and their relation to the effects of individual mutations.}
The top panel shows the subset of the full H3N2 HA tree in Figure~\ref{suppfig:tree} windowed from 2004 to 2014. 
Circles indicate amino-acid mutations, and are colored according to the mutational effect measured in our deep mutational scanning (negative values indicate mutations measured to be deleterious to viral growth).
The bottom panel shows the frequency trajectory of the clade downstream of each mutation, with clades colored according to the mutational effects as in the top panel.
It is obvious that most clades that reach high frequency have mutations that are measured to be relatively favorable in our experiments.
\comment{John: Can you update figure? We defer to you and Trevor on choice of colors, I know there was some Slack discussion of this but can't remember answer. But definitely let's increase the size of the legend and make it $<-10$, and $> 0$ rather than having the \emph{inf} in the labels. Also, can you make y-axis at bottom ``clade frequency'' rather than ``mutation frequency'' as we use that terminology in the text instead.}
}
\end{figure}

@@

\begin{figure*}[ht]
\centering
\includegraphics[width=\textwidth]{figs/muteffect_by_maxfreq.pdf}
\caption{\label{fig:muteffect_maxfreq}
{\bf The experimental measurements are informative about the evolutionary fate of viral clades.}
(A) The correlation between the effects of mutations as measured in our experiments and the maximum frequency reached by viral clades.
The plots show the Spearman correlation $\rho$ and a $P$-value obtained by repeatedly randomizing the experimental measurements among sites \comment{John: is this right about how the $P$-value was calculated?}.
(B) The distribution of mutational effects among clades binned by the maximum frequency that they reach.
Note that the analysis is performed separately for the pre- and post-Perth/2009 partitions of the tree (see Figure~\ref{suppfig:tree}).
\comment{John: label the top and bottom into A and B as indicated in the legend above, adjust font sizes (particularly, tick labels too big), adjust capitalization to be consistent with other figures (which usually start lower-case unless proper noun), increasing spacing between A and B panels. Also, no need for a title at top and figure can probably get smaller. Some of this might be best done in Inkscape working with Juhye.}
}
\end{figure*}

@@

\begin{figure*}[ht]
\centering
\includegraphics[width=\textwidth]{figs/muteffect_by_maxfreq_WSN_1.pdf}
\caption{\label{fig:muteffect_maxfreq_WSN}
{\bf Experimental measurements on an H1 HA are not informative about the evolutionary fate of H3N2 viral clades.}
This figure repeats the analysis of the H3N2 clade frequencies in Figure~\ref{fig:muteffect_maxfreq}A, but uses the deep mutational scanning data for an H1 HA as measured in \cite{doud2016accurate}.
Figure~\ref{suppfig:muteffect_maxfreq_WSN_supp} shows the histograms comparable to those in Figure~\ref{fig:muteffect_maxfreq}B.
\comment{John: Can you match any formatting changes that you make to Figure~\ref{fig:muteffect_maxfreq}A here so that they look consistent.}
}
\end{figure*}

@@

\begin{suppfigure}[H]
\centerline{\includegraphics[width=0.75\textwidth]{figs/muteffect_by_maxfreq_WSN_2.pdf}}
\caption{\label{suppfig:muteffect_maxfreq_WSN_supp}
{\bf The distribution of mutational effects measured in H1 HA among H3N2 clades binned by the maximum frequency that they reach.}
This figure repeats the analysis of the H3N2 clade frequencies in Figure~\ref{fig:muteffect_maxfreq}B, but uses the deep mutational scanning data for an H1 HA as measured in \cite{doud2016accurate}.
\comment{John: Can you match any formatting changes that you make to Figure~\ref{fig:muteffect_maxfreq}B here so that they look consistent.}
}
\end{suppfigure}
